Quantitative Benchmarking for Automatically Generated Conjectures

### 2017-11-27 TEBenchmarkPaper/article.tex
More text

\begin{figure}
  \centering
  \input{time.pgf}
  \caption{Running times of QuickSpec on theories sampled from TIP. Each point
    is an individual run, with 30 per size except for 1 (14) and 2 (29). Colour
    indicates the total number of conjectures produced by each run, with failing
    processes shown in red. Some points are shifted horizontally to prevent
    overlaps. Standard box plots are included, although many are obscured due to
    quartiles near zero or timeout.}
  \label{figure:quickspec_results}
\end{figure}

\begin{figure}
  \centering
  \input{prec.pgf}
  \caption{}
  \label{figure:quickspec_results}
\end{figure}

We applied our sampling methodology to the 269 TIP definitions for sample sizes
between 1 and 20. We chose 30 samples for each size, but due to the requirement
for samples to admit at least one theorem, only 14 unique samples were obtained
for size 1 and 29 for size 2. A timeout of 3 minutes was imposed, since
preliminary experiments showed that most runs (around $\frac{4}{5}$) either
finish before this time, or still do not finish after an hour. The results are
shown in figure~\ref{figure:quickspec_results}.

The only errors we encountered were timeouts, and these occurred for all sample
sizes. Whilst the median runtimes are mostly in the range of a few seconds, the
variance increased dramatically towards size 10, and $\frac{1}{4}$ of sample
size 13 timed out.

Most runs finished quickly and generated few conjectures, with the most being
131 conjectures from a sample of size 20. Sizes 15 to 20 show a slight but
consistent reduction in runtimes, which is likely an artefact of our testing
setup.

Precision and recall were calculated by normalising each equation generated by
QuickSpec: all variables are identified numerically, in ascending order from
left to right; equations were also sorted such that the expression on the
left-hand-side is lexicographically smaller than that on the right.

Theorem statements from TIP were first updated to reference only the 269
distinct names we explored, and references to constructors and destructors were
updated to use the $\eta$-expanded versions instead. Equations were normalised
in the same way as above, whilst non-equations and those containing
\texttt{lambda} functions (which QuickSpec cannot synthesise) were replaced with
null values. Elements of these two sets were deemed to match if they were
syntactically identical.

@@

If ATE matures to the point of providing \emph{novel} insights, even in
well-studied domains, our assumption breaks down because such results cannot (by
definition) appear in any existing corpus.

@@

\subsection{Technical Difficulties}

Our benchmarking approach only applies to systems which act in ``batch mode'',
i.e. those which choose to halt after emitting some output. Whilst all of the
systems we have encountered are of this form, an analogous benchmark may be
desirable for systems which instead run continuously or interactively (perhaps
akin to McCarthy's ``advice taker''%FIXME~\cite{}.

@@

Our benchmark suite provides a unifying goal for the diverse approaches of
theory exploration, whilst avoiding some of the philosophical complications of
the field. Whilst the current implementation is quite limited, we welcome
additions from other researchers, to more closely align the benchmark with their
goals, and the strenghts and weaknesses of their systems.

``Solving'' this benchmark suite would not solve the problem of theory
exploration in general, so more ambitious goals must be set in the future; but
we have found that there is still a long way to go until that problem arises.

>>>>

\begin{figure}
  \centering
  \input{time.pgf}
  \caption{Running times of QuickSpec on theories sampled from TIP. Each point
    is an individual run, with 30 per size except for 1 (14) and 2 (29). Colour
    indicates the total number of conjectures produced by each run, with failing
    processes shown in red. Some points are shifted horizontally to prevent
    overlaps. Standard box plots are included, although many are obscured due to
    quartiles near zero or timeout. Whiskers show points at 1.5 $\times$ the
    inter-quartile range.}
  \label{figure:quickspec_runtimes}
\end{figure}

\begin{figure}
  \centering
  \input{prec.pgf}
  \caption{Precision and recall of all runs (spread horizontally to avoid
    overlaps). Point colour shows how many conjectures were in the ground truth
    set for each run. Lines show the mean value (average of ratios) and the
    shaded region is standard deviation, assuming a binomial distribution.}
  \label{figure:quickspec_precRec}
\end{figure}

We applied our sampling methodology to the 269 TIP definitions for sample sizes
between 1 and 20. We chose 30 samples for each size, but due to the requirement
for samples to admit at least one theorem, only 14 unique samples were obtained
for size 1 and 29 for size 2. A timeout of 3 minutes was imposed, since
preliminary experiments showed that most runs either finish before this time, or
still do not finish after an hour. The running times are shown in
figure~\ref{figure:quickspec_runtimes}.

The only errors we encountered were timeouts (shown in red), and these occurred
for all sample sizes. Whilst the median runtimes are mostly in the range of a
few seconds, the variance shows two distinct regions: sizes below 10 are mostly
finished within 30 seconds, whilst sizes 10 and above have large variance; the
whole upper quartile of size 13 timed out. There is also an unexpected downwards
trend towards size 20, and a slight but consistent reduction in minimum runtimes
from sizes 15 to 20 which is likely an artefact of our testing setup.

Most runs finished quickly and generated few conjectures, with the most being
131 conjectures from a sample of size 20.

Equations generated by QuickSpec and those from the ground truth were normalised
by numbering their variables and putting the left- and right-hand-sides in
lexicographical order. These two sets were then compared syntactically for
precision and recall, with the results shown in
figure~\ref{figure:quickspec_precRec}.

We use a binomial distribution to model precision and recall, as if each
conjecture is a ball drawn from an urn of wanted/unwanted results (in the case
of precision) or found/missed ground truths (in the case of recall). The mean of
these distributions estimates the overall proportion of wanted or found
conjectures, and we use the standard deviation to determine how well this mean
fits the data.

There is a clear negative trend in precision as the sample size increases,
indicating that a more aggressive post-processing filter may be useful as we
explore larger theories. Some small samples reached full precision, but nothing
above size 10 exceeded 50\%.

Recall shows a slight downward trend but is mostly flat, averaging between 10\%
and 40\%. This is expected, since adding more definitions to a QuickSpec
signature will never reduce the conjectures it will find. Those conjectures
which QuickSpec failed to find include terms which QuickSpec cannot synthesise
(such as conditional equations, inequalities, terms larger than the search depth
and those with anonymous functions) as well as equations which are removed as
``uninteresting'' by the final congruence closure step.

@@

If ATE matures to the point of providing \emph{truly novel} insights, even in
well-studied domains, our assumption breaks down because such results cannot (by
definition) appear in any existing corpus.

A fundamental limitation of our benchmarking approach is that it only applies to
systems which act in ``batch mode'', i.e. those which choose to halt after
emitting some output. Whilst all of the systems we have encountered are of this
form, an analogous benchmark may be desirable for systems which instead run
continuously or interactively (perhaps akin to McCarthy's ``advice taker'').%FIXME~\cite{}.

@@

@@

Our benchmarking methodology, and accompanying suite based on TIP, provides a
general way to evaluate and compare many diverse approaches to the problem of
theory exploration, whilst avoiding some of the philosophical complications of
the field. The methodology can be adapted to suit particular domains, for
example our experiments have focused on inductive theories with higher-order
functions, and the corpora used to generate a benchmark suite can always be
improved and refined with the addition of new definitions and theorems.

We believe that a standard approach to benchmarking and comparison will ease the
burden on researchers wanting to evaluate systems, and provide a common goal to
pursue in the short term.

``Solving'' this benchmark suite would not solve the problem of theory
exploration in general, so more ambitious goals must be set in the future; but
we believe that our approach will provide a compelling challenge, at least for
the foreseeable future.
